\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Estimating Trending Topics on Twitter with Small Subsets of the Total Data}

### 2015-01-13 paper.tex
Minor paper edits

Our approach was to explore whether or not adding historical data would improve the accuracy of 
trending hashtag detection. To that end, we implemented two algorithms: one called the naive algorithm, the other called the history-sensitive algorithm. We wanted to test whether an algorithm that uses
the history (in condensed form) to help understand what is trending at the current time could outperform an algorithm with a very short term memory.

@@

Now we analyze the time due to the end-of-day updates to the History and the resulting updates to the heap. By Lemma $5$ of the Hokusai paper, the amortized time required to do all end-of-day aggregation is $O(md)$. Then, for each of the $s$ nodes in the heap, it takes $O(d)$ time to compute each updated key and $O(\log(s))$ amortized time to update the heap given the new key.

Thus, it takes $O(md + slog(s))$ amortized time $+ O(ds)$ time to do all necessary updates at the end of the day.

Querying the data structure for the top $k$ trending items takes $O(k*log(s))$ amortized time for $\textbf{delete-max}$ operations, $O(k)$ time to announce that these items are trending, and $O(k)$ amortized time to reinsert these nodes into the heap.

Thus, it takes $O(k*log(s))$ amortized time to determine what is trending.

@@

The History requires $O(md)$ space for each Count-Min sketch, so it requires a total of $O(md*\log(T))$ space.
Each node in the heap requires a constant amount of space, so the heap requires $O(s)$ space.
The hash table always contains at most $s$ entries with each entry requiring a constant amount of space.  Also, in order to maintain an expected $O(1)$ lookup time, the hash table needs to have $O(s)$ bins.  Thus, the hash table requires $O(s)$ space.
The queue requires an entry for every hashtag seen in the last $y$-unit, so it requires $O(x)$ space.

Thus, everything requires $O(md*\log(T) + x)$ space since $s < x$.

@@

% y is 3 hours, z is 1 day
\subsection{Choosing the parameters $y$ and $z$}
For $y$, the unit time period we attempted to find the trending hashtags for, we chose a $3$-hour interval.
For $z$, the unit time period for the History data structure, we chose day-length intervals. These values made sense because the testing data we used occurred over a $30$-day interval, and we wanted to avoid too fine a temporal resolution so that we could run our code on the data in reasonable time.

@@

where the error in the query is in factor of $epsilon$ with probability $1 - \delta$. By Theorem $1$ in \cc{Matusevych:2012}, we have that if $n_x$ is the true count of a hashtag stored in the CM sketch, and $c_x$ is the estimate, we are guaranteed $n_x \leq c_x \leq n_x + \epsilon \sum_{x'} n_{x'}$. 
By choosing $m = 3500$, we have that $\epsilon = \frac{e}{3500}$ and $1 - \delta = 1 - \frac{1}{e^{20}}$, which is to say that we are within that error with very high probability. The error term is a multiplier on the total number of hashtags, of which there are $\sim 15$ million, as an upper bound. Thus we are within very high probability of the true count with range $\sim 12000$, on a given count. As a result, by storing less memory and keeping the memory down more, we get some amount of smoothing for free.

@@

We downloaded the \href{https://archive.org/details/twitterstream}{September $2014$ archive of Twitter data } to serve as test data for our algorithms \cc{Twitter2014}. We then parsed the data into (timestamp, hashtag) format
for all thirty days worth of data. Using these files as pseudo-real-time input, we used the last week of the data as a benchmark for comparison between the naive and the history-sensitive algorithms. We divided the data into three-hour chunks, and queried for the top ten hashtags in each of the three-hour time blocks
in the last week of the data using both algorithms. Note that the history-senstive algorithm was required to parse through the first three weeks as well in order to build its history, whereas the naive algorithm was only required to pay attention to six-hour chunks at a time, and thus never needed to look a bit beyond the scope of that week.

@@

We plot it below in the histogram shown in \aref{fig:jaccard_sims}. The $x$-axis is the Jaccard similarity, and the $y$-axis is the count. The mean was $0.47$, and the standard Deviation was $0.18$. From the plot, we can see that the distribution of the Jaccard similarities is roughly a centered normal distribution around $0.5$. A Jaccard similarity of $0.5$ means about half the hashtags are the same in each $3$-hour time interval between the two algorithms.

@@

In \aref{fig:news_vis}, we have superimposed pictures from 
news articles that the tweets describe on a time axis, as
the hashtag referencing the article starts trending. The Jaccard similarities 
for each of these three-hour intervals were all above $0.5$, so both algorithms
found the hashtags that we graph in the figure.
We checked that the news was released (via Yahoo! News, Twitter, Google News, ESPN, and others)
at around the time our algorithms say the hashtag relating to the news
begins to trend. A fair number of the trending hashtags related to sports and TV shows.

@@

Based on the approximately normal, centered distribution of the Jaccard similarities, we observe that history-sensitive performed about as well as the naive approach in terms of getting trending tweets. However, the naive algorithm uses considerably less space and time (since we only need to check the 
$y$-unit before the current $y$-unit) to get similar results. Thus we conclude that the space efficiency of the naive approach outweighs the benefits of the history-sensitive approach for this particular task. 
is likely superior to the history-sensitive approach. Furthermore, in our development of the history-sensitive algorithm, utilized larger errors and smoothing to perform decently, suggesting that
our approach to evaluating the history may not be the best one for tweet-trending detection.

@@

Furthermore, for real data from all of the Twitter firehose, we would require a better hardware infrastructure dedicated to processing all the data. Ideally we would use distributed systems and would
parallelize our code before running it. The fact that the code is broken up into disparate parts make it easier to implement in parallel.

@@

We also know that an estimation the top $k$ hashtags in a given time interval could provide a topic model for tweets. It could be interesting to use the top $k$ hashtags to build a time-sensitive topic model and analyze quantities including the speed of topic change, the variety of topics, and so on.

Even though not perhaps irrelevant to the specific task of identifying trending hashtags, 
the approach of the history-sensitive algorithm may be useful in other domains.
Future work could also investigate the performance of the history-sensitive algorithm on other data streams of interest, for instance, Wikipedia edits -- our goal would be to estimate which Wikipedia topics are being edited the most at any given time interval of some length. Another interesting application of this algorithm to test would be the real-time estimatation of the hottest selling stocks on Wall Street to see if either of these settings results in a significant improvement over the naive algorithm.

>>>>

Our approach was to explore whether or not adding historical data would improve the accuracy of 
trending hashtag detection. To that end, we implemented two algorithms: one called the naive algorithm, the other called the history-sensitive algorithm. We wanted to test whether an algorithm that uses
the history (in condensed form) to better understand what is trending at the current time could outperform an algorithm with a very short term memory.

@@

Now we analyze the time due to the end-of-day updates to the History and the resulting updates to the heap. By Lemma $5$ in \cc{Matusevych:2012}, the amortized time required to do all end-of-day aggregation is $O(md)$. Then, for each of the $s$ nodes in the heap, it takes $O(d)$ time to compute each updated key and $O(\log(s))$ amortized time to update the heap given the new key.

Thus, it takes $O(md + s\log(s))$ amortized time $+ O(ds)$ time to do all necessary updates at the end of the day.

Querying the data structure for the top $k$ trending items takes $O(k\log(s))$ amortized time for $\textbf{delete-max}$ operations, $O(k)$ time to announce that these items are trending, and $O(k)$ amortized time to reinsert these nodes into the heap.

Thus, it takes $O(k\log(s))$ amortized time to determine what is trending.

@@

The History requires $O(md)$ space for each Count-Min sketch, so it requires a total of $O(md\log(T))$ space.
Each node in the heap requires a constant amount of space, so the heap requires $O(s)$ space.
The hash table always contains at most $s$ entries with each entry requiring a constant amount of space.  Also, in order to maintain an expected $O(1)$ lookup time, the hash table needs to have $O(s)$ bins.  Thus, the hash table requires $O(s)$ space.
The queue requires an entry for every hashtag seen in the last $y$-unit, so it requires $O(x)$ space.

Thus, everything requires $O(md\log(T) + x)$ space since $s < x$.

@@

% y is 3 hours, z is 1 day
\subsection{Choosing the parameters $y$ and $z$}
For $y$, the unit time period we attempted to find the trending hashtags for, we chose a $3$-hour interval.
For $z$, the unit time period for the History data structure, we chose day-length intervals. These values made sense because the testing data we used occurred over a $30$-day interval, and we wanted to avoid too fine a temporal resolution so that we could run our code on the data in a reasonable amount of time.

@@

where the error in the query is in factor of $\epsilon$ with probability $1 - \delta$. By Theorem $1$ in \cc{Matusevych:2012}, we have that if $n_x$ is the true count of a hashtag stored in the CM sketch, and $c_x$ is the estimate, we are guaranteed $n_x \leq c_x \leq n_x + \epsilon \sum_{x'} n_{x'}$. 
By choosing $m = 3500$, we have that $\epsilon = \frac{e}{3500}$ and $1 - \delta = 1 - \frac{1}{e^{20}}$, which is to say that we are within that error with very high probability. The error term is a multiplier on the total number of hashtags, of which there are $\sim 15$ million, as an upper bound. Thus we are within very high probability of the true count with range $\sim 12000$, on a given count.

By using less memory we increase the amount of error.  These error terms are added to the denominator of our priority function, and as a result, they prevent hashtags with very relatively low frequency from being identified as trending, which is a very desirable property.

@@

We downloaded the \href{https://archive.org/details/twitterstream}{September $2014$ archive of Twitter data } to serve as test data for our algorithms \cc{Twitter2014}. We then parsed the data into (timestamp, hashtag) pairs
for all thirty days worth of data. Using these files as pseudo-real-time input, we used the last week of the data as a benchmark for comparison between the naive and the history-sensitive algorithms. We divided the data into three-hour chunks, and queried for the top ten hashtags in each of the three-hour time blocks
in the last week of the data using both algorithms. Note that the history-senstive algorithm was required to parse through the first three weeks as well in order to build its history, whereas the naive algorithm was only required to pay attention to six-hour chunks at a time, and thus never needed to look a bit beyond the scope of that week.

@@

We plot this data in the histogram shown in \aref{fig:jaccard_sims}. The $x$-axis is the Jaccard similarity, and the $y$-axis is the count. The mean was $0.47$, and the standard Deviation was $0.18$. From the plot, we can see that the distribution of the Jaccard similarities is roughly a centered normal distribution around $0.5$. A Jaccard similarity of $0.5$ means about half the hashtags are the same in each $3$-hour time interval between the two algorithms.

@@

In \aref{fig:news_vis}, we have superimposed pictures from 
news articles that the tweets describe on a time axis, as
the hashtag referencing the article starts trending. The Jaccard similarities 
for each of these three-hour intervals were all above $0.5$, so both algorithms
found the hashtags that we graph in the figure.
We checked that the news was released (via Yahoo! News, Twitter, Google News, ESPN, and others)
at around the time our algorithms say the hashtag relating to the news
begins to trend. A fair number of the trending hashtags are related to sports and TV shows.

@@

Based on the approximately normal, centered distribution of the Jaccard similarities, we observe that the history-sensitive approach performed about as well as the naive approach in terms of getting trending tweets. However, the naive algorithm uses considerably less space and time (since we only need to check the 
$y$-unit before the current $y$-unit) to get similar results. Thus we conclude that the space efficiency of the naive approach outweighs the benefits of the history-sensitive approach for this particular task.  Furthermore, in our development of the history-sensitive algorithm, we utilized larger errors as a method of smoothing to perform decently, suggesting that our approach for evaluating the history may not be the best one for tweet-trending detection.

@@

Furthermore, for real data from all of the Twitter firehose, we would require a better hardware infrastructure dedicated to processing all the data. Ideally we would use distributed systems and would
parallelize our code before running it. The fact that the code is broken up into disparate parts makes it easier to implement in parallel.

@@

We also know that an estimation of the top $k$ hashtags in a given time interval could provide a topic model for tweets. It could be interesting to use the top $k$ hashtags to build a time-sensitive topic model and analyze quantities including the speed of topic change, the variety of topics, and so on.

Even though it is perhaps not suited for the specific task of identifying trending hashtags, 
the approach of the history-sensitive algorithm may be useful in other domains.
Future work could also investigate the performance of the history-sensitive algorithm on other data streams of interest, for instance, Wikipedia edits -- our goal would be to estimate which Wikipedia topics are being edited the most at any given time interval of some length. Another interesting application of this algorithm to test would be the real-time estimatation of the hottest selling stocks on Wall Street to see if either of these settings results in a significant improvement over the naive algorithm.
